---
output:
  html_document: default
  pdf_document: default
---


# Data Dictionary

Variable	Definition	Key
survival	Survival	0 = No, 1 = Yes
pclass	Ticket class	1 = 1st, 2 = 2nd, 3 = 3rd
sex	Sex	
Age	Age in years	
sibsp	# of siblings / spouses aboard the Titanic	
parch	# of parents / children aboard the Titanic	
ticket	Ticket number	
fare	Passenger fare	
cabin	Cabin number	
embarked	Port of Embarkation	C = Cherbourg, Q = Queenstown, S = Southampton
Variable Notes
pclass: A proxy for socio-economic status (SES)
1st = Upper
2nd = Middle
3rd = Lower

age: Age is fractional if less than 1. If the age is estimated, is it in the form of xx.5

sibsp: The dataset defines family relations in this way...
Sibling = brother, sister, stepbrother, stepsister
Spouse = husband, wife (mistresses and fiancés were ignored)

parch: The dataset defines family relations in this way...
Parent = mother, father
Child = daughter, son, stepdaughter, stepson
Some children travelled only with a nanny, therefore parch=0 for them


## GitHub Dataset Link

[GitHub](https://github.com/AkshayMogha/Titanic.git)


```{r}
#setting working directory

setwd("C:/Users/Akshay Rajput/Downloads/titanic")
#loading raw data 
train<-read.csv("train.csv",header=TRUE,stringsAsFactors = FALSE)
test<-read.csv("test.csv",header=TRUE,stringsAsFactors = FALSE)
#adding a survived variable to test
test_survived<-data.frame(Survived=rep("None",nrow(test)),test[,])

#combining train and test dataset
titanic<-rbind(train,test_survived)

```
# Structure of Titanic Dataset
```{r,echo=FALSE}
str(titanic)
```

**Checking for duplicate names in titanic having train and test**
```{r}
duplicate.names<-as.character(titanic[which(duplicated(as.character(titanic$Name))),"Name"])
```

```{r,echo=FALSE}
duplicate.names
```
**checking whether we have same observations again**


```{r}
titanic[which(titanic$Name %in% duplicate.names ),]
```

**No, Passengers have different PassengerId**

**Checking and Filling NA values in Variables**

```{r}
which(is.na(titanic$PassengerId))
```
```{r}
which(is.na(titanic$Survived))
```
```{r}
which(is.na(titanic$Pclass))
```
```{r}
which(is.na(titanic$Name))
```
```{r}
which(is.na(titanic$Sex))
```
```{r}
which(is.na(titanic$Age))
```
```{r}
which(is.na(titanic$SibSp))
```
```{r}
which(is.na(titanic$Parch))
```
```{r}
which(is.na(titanic$Ticket))
```
```{r}
which(is.na(titanic$Fare))
```
```{r}
which(is.na(titanic$Cabin))
```
```{r}
which(is.na(titanic$Embarked))
```


**We  have two variables with na values i.e age and fare**

# For fare
```{r}
FareNA<-mean(titanic[which(!is.na(titanic$Fare)),"Fare"])
titanic[which(is.na(titanic$Fare)),"Fare"]<-FareNA
```



```{r}
library(stringr)
```

# For age

```{r}
titanicNA<-titanic[which(!is.na(titanic$Age)),]

miss<-titanicNA[which(str_detect(titanicNA$Name,"Miss.")),]
missNA<-round(mean(miss$Age))
```
**Mean age of Misses. to replace NAs in Age variable for Miss. titles**
```{r,echo=FALSE}
missNA
```

```{r}
mrs<-titanicNA[which(str_detect(titanicNA$Name,"Mrs.")),]
mrsNA<-round(mean(mrs$Age))
```
**Mean age of Mrs. to replace NAs in Age variable for Mrs. titles**
```{r,echo=FALSE}
mrsNA
```

```{r}
master<-titanicNA[which(str_detect(titanicNA$Name,"Master.")),]
masterNA<-round(mean(master$Age))
```
**Mean age of Master. to replace NAs in Age variable for Master. titles**
```{r,echo=FALSE}
masterNA
```

```{r}
mr<-titanicNA[which(str_detect(titanicNA$Name,"Mr.")),]
mrNA<-round(mean(mr$Age))
```
**Mean age of Mr. to replace NAs in Age variable for Mr. titles**
```{r,echo=FALSE}
mrNA
```


```{r}
NAage<-titanic[which(is.na(titanic$Age)),]

NAagemiss<-NAage[which(str_detect(NAage$Name,"Miss.")),"PassengerId"]
NAagemrs<-NAage[which(str_detect(NAage$Name,"Mrs.")),"PassengerId"]
NAagemaster<-NAage[which(str_detect(NAage$Name,"Master.")),"PassengerId"]
NAagemr<-NAage[which(str_detect(NAage$Name,"Mr.")),"PassengerId"]


for(i in 1:length(NAagemiss)){
  titanic[NAagemiss[i],"Age"]<-missNA}
for(i in 1:length(NAagemrs)){
  titanic[NAagemrs[i],"Age"]<-mrsNA}
for(i in 1:length(NAagemaster)){
  titanic[NAagemaster[i],"Age"]<-masterNA}
for(i in 1:length(NAagemr)){
  titanic[NAagemr[i],"Age"]<-mrNA}

NAextramean<-round(mean(titanicNA$Age))
NAextra<-titanic[which(is.na(titanic$Age)),"PassengerId"]
```
**Mean value of all ages to be filled in NAs which has no titles**
```{r,echo=FALSE}
NAextramean
```
```{r}
for(i in 1:length(NAextra))
{titanic[NAextra[i],"Age"]<-NAextramean}
```

**Now there is no NA values . Data is clean**


# Checking for blank values

**Replacing Blank values in Embarked with S**

```{r}
for(i in 1:nrow(titanic))
{if(titanic[i,"Embarked"]!="S"&titanic[i,"Embarked"]!="C"&titanic[i,"Embarked"]!="Q")
{
  titanic[i,"Embarked"]<-"S"
  
  
}
  
}
```

# Distribution of data

# BOX PLOT
```{r}
boxplot(titanic$Age~titanic$Sex)
```
```{r}
boxplot(titanic$Fare~titanic$Age)
```
```{r}
boxplot(titanic$Fare~titanic$Sex)
```
```{r}
boxplot(titanic$Fare~titanic$Embarked)
```

# DENSITY PLOT
```{r}
plot(density(titanic$Fare),
     col="red",
     main="Density plot for Fare",
     xlab = "Fare",
     ylab="Dnesity")
polygon(density(titanic$Fare),col = "red")
```

## SCATTER PLOT
```{r}
plot(titanic$Age, titanic$Fare, main = "flow",
     xlab = "Age", ylab ="Fare",
     pch = 19, frame = FALSE)
```

 **Adding regression line**
```{r}
plot(titanic$Age, titanic$Fare, main = "flow",
     xlab = "Age", ylab = "Fare",
     pch = 19, frame = FALSE)
abline(lm(Fare ~ Age, data = titanic), col = "blue")
```
## It shows Fare is increasing with Age





**Converting Survived and Pclass variables into factors**
```{r}
titanic$Survived<-as.factor(titanic$Survived)
titanic$Pclass<-as.factor(titanic$Pclass)
```


**count of factor classes of each variable factor**
```{r,echo=FALSE}
table(titanic$Survived)
table(titanic$Pclass)
```

# Hypothesis - Pclass 1 passengers are more in survived numbers
# cause they are rich and had paid more 

```{r,echo=FALSE}
outof<-"survived out of"
percentage<-"%"
```
```{r}
survived.1<-length(train[which(train$Survived==1 & train$Pclass==1),"PassengerId"])
Pclass1<-length(train[which(train$Pclass==1),"PassengerId"])
survived.1.per<-(survived.1/Pclass1)*100
```
**No. of passangers which survived and belongs to Pclass 1**

```{r,echo=FALSE}
paste(survived.1,outof,Pclass1,collapse = " ")
```
```{r,echo=FALSE}
paste(survived.1.per,percentage,collapse=" ") 
```

```{r}
survived.2<-length(train[which(train$Survived==1 & train$Pclass==2),"PassengerId"])
Pclass2<-length(train[which(train$Pclass==2),"PassengerId"])
survived.2.per<-(survived.2/Pclass2)*100
```
**No. of passangers which survived and belongs to Pclass 2**

```{r,echo=FALSE}
paste(survived.2,outof,Pclass2,collapse = " ")
```
```{r,echo=FALSE}
paste(survived.2.per,percentage,collapse=" ") 
```

```{r}
survived.3<-length(train[which(train$Survived==1 & train$Pclass==3),"PassengerId"])
Pclass3<-length(train[which(train$Pclass==3),"PassengerId"])
survived.3.per<-(survived.3/Pclass3)*100
```
**No. of passangers which survived and belongs to Pclass 3**
```{r,echo=FALSE}
paste(survived.3,outof,Pclass3,collapse = " ")
```
```{r,echo=FALSE}
paste(survived.3.per,percentage,collapse=" ") 
```

```{r}
total.survived<-sum(survived.1,survived.2,survived.3)
PassengerId.total<-length(train$PassengerId)
total.survived.per<-(total.survived/PassengerId.total)*100
```
**Percentage of passengers total survived**

```{r,echo=FALSE}
paste(total.survived.per,percentage,collapse = " ")
```

```{r}
library(plotrix)

pct=round(table(train$Survived)/sum(table(train$Survived))*100)
lbs=paste(c("Dead","Survived")," ",pct,"%",sep=" ")
```

**Plotting Pie Chart for survived and Dead**

```{r,echo=FALSE}
pie3D(table(train$Survived),labels = lbs,main="Pie chart showing percentag of survived vs dead")
```


```{r}
piesurvived1<-c(survived1=survived.1,dead1=Pclass1-survived.1)
pct=round(piesurvived1/sum(piesurvived1)*100)
lbs=paste(c("Survived","Dead")," ",pct,"%",sep=" ")
```
**Plotting Pie Chart for survived.1 for Pclass1 or not**

```{r,echo=FALSE}
pie3D(piesurvived1,labels = lbs,main="Pie chart showing percentag of survived vs dead of Pclass 1")
```

```{r}
piesurvived2<-c(survived2=survived.2,dead2=Pclass2-survived.2)
pct=round(piesurvived2/sum(piesurvived2)*100)
lbs=paste(c("Survived","Dead")," ",pct,"%",sep=" ")
```
**Plotting Pie Chart for survived.2 for Pclass2 or not**

```{r,echo=FALSE}
pie3D(piesurvived2,labels = lbs,main="Pie chart showing percentag of survived vs dead of Pclass 2")
```

```{r}
piesurvived3<-c(survived3=survived.3,dead3=Pclass3-survived.3)
pct=round(piesurvived3/sum(piesurvived3)*100)
lbs=paste(c("Survived","Dead")," ",pct,"%",sep=" ")
```
**Plotting Pie Chart for survived.3 for Pclass3 or not**

```{r,echo=FALSE}
pie3D(piesurvived3,labels = lbs,main="Pie chart showing percentag of survived vs dead of Pclass 3")
```

```{r}
library(ggplot2)
```
**Plotting bar chart for survival of each pclass**

```{r,echo=FALSE}
ggplot(train,aes(x=Pclass,fill=factor(Survived)))+
  geom_bar(position="dodge")+
  xlab("Pclass")+
  ylab("No. of Passengers survived/not survived")+
  labs(fill="0 : Dead / 1 : Survived")
```
```{r}
library(scales)
library(dplyr)
```
```{r}
train$Survived=as.factor(train$Survived)
train$Pclass=as.factor(train$Pclass)
```
**Structure of train dataset after factoring Survived and Pclass variables**

```{r,echo=FALSE}
str(train)
```
**Forming a new dataset using pipe %>% operator**
```{r}
Pclass_survival<- train%>%group_by(Pclass,Survived)%>%dplyr::summarize(No._of_Passengers_survived_not_survived=n())
```
**plotting the survival based on each Pclass but using new dataframe created using %>% operator**

```{r,echo=FALSE}
ggplot(Pclass_survival,aes(Pclass,No._of_Passengers_survived_not_survived,fill=Survived))+
  geom_bar(position="dodge",stat="identity")+
  ggtitle("bar chart for survived in each pclass")+
  scale_y_continuous(labels = comma)+
  scale_fill_manual(values = c("#CC0000", "#006600"))
```
  

**Factoring Sex variable**

```{r}
titanic$Sex<-as.factor(titanic$Sex)
```

**examining the survival with gender bar chart**

```{r,echo=FALSE}
ggplot(train,aes(Sex,fill=Survived))+
  geom_bar(position = "dodge")+
  ggtitle("survival with Sex")+
  xlab("Sex")+
  ylab("Total Survived")+
  labs(fill="Survived")
```
  
```{r}
female.sur<-train%>%select(Sex,Survived)%>%filter(Sex=="female" & Survived==1)
female.survive<-length(female.sur$Survived)

female.de<-train%>%select(Sex,Survived)%>%filter(Sex=="female" & Survived==0)
female.dead<-length(female.de$Survived)

femalesurvived<-c(femalesur=female.survive,femaledead=female.dead)
pct=round(femalesurvived/sum(femalesurvived)*100)
lbs=paste(c("Survived","Dead")," ",pct,"%",sep=" ")
```
**Examining the survival with female with pie chart**

```{r,echo=FALSE}
pie3D(femalesurvived,labels = lbs,main="female with Survival")
```

```{r}
male.sur<-train%>%select(Sex,Survived)%>%filter(Sex=="male" & Survived==1)
male.survive<-length(male.sur$Survived)

male.de<-train%>%select(Sex,Survived)%>%filter(Sex=="male" & Survived==0)
male.dead<-length(male.de$Survived)

malesurvived<-c(malesur=male.survive,maledead=male.dead)
pct=round(malesurvived/sum(malesurvived)*100)
lbs=paste(c("Survived","Dead")," ",pct,"%",sep=" ")
```
**Examining the survival with male with pie chart**

```{r,echo=FALSE}
pie3D(malesurvived,labels = lbs,main="male with Survival")
```

```{r}
sexsurvive<-c(female.survive,male.survive)
pct<-round(sexsurvive/(sum(malesurvived)+sum(femalesurvived))*100)
lbs=paste(c("Female","Male")," ",pct,"%",sep=" ")
```
**Examining survival with males and females**

```{r,echo=FALSE}
pie3D(sexsurvive,labels=lbs,main="Survival per Sex")
```



# Examining the names

```{r}
head(as.character(train$Name))
```


**examining survival rate by TITLES such as Mr. ,Mrs. ,etc**
**first we have to make a title variable**

```{r}
title<-NULL
```

**We made titler function to extract titles from names**
```{r}
titler<-function(Name)
{
  name<-as.character(Name)
  if(length(grep("Miss.",name))>0)
  {return ("Miss.")}
  else if(length(grep("Mrs.",name))>0)
  {return("Mrs.")}else if(length(grep("Master.",name))>0)
  {return("Master.")}else if(length(grep("Mr.",name))>0)
  {return("Mr.")}
  else{
    for(i in 1:nrow(titanic)){
    if(titanic[i,"Name"]==name)
    {if(titanic[i,"Sex"]=="male")
    {if(titanic[i,"Age"]<=masterNA)
    {return("Master.")}
      else{return("Mr.")}
      
    }
      else{
        if(titanic[i,"Age"]<=sum(missNA,mrsNA)/2)
        {return("Miss.")}
        else{return("Mrs.")}
      }
      break
    }}}}
```

**Assigning values and Adding title variable as a factor to titanic dataset**

```{r}
for(i in 1:nrow(titanic))
{title<-c(title,titler(titanic[i,"Name"]))}

titanic$title<-as.factor(title)
```

**Now we will plot the survival based on titles using train set only,**
**but we have made changes to titanic set so we will work with only upper observations**
**which comes from train dataset**

```{r,echo=FALSE}
ggplot(titanic[1:nrow(train),],aes(x=title,fill=Survived))+
  geom_bar(position="dodge")+
  # facet_wrap(~Pclass)+
  ggtitle("titlewise survival")+
  xlab("title")+
  ylab("No. of people survived/dead")+
  labs(fill="Survived")
```


**Division on Pclass**

```{r,echo=FALSE}
ggplot(titanic[1:nrow(train),],aes(x=title,fill=Survived))+
  geom_bar(position="dodge")+
  facet_wrap(~Pclass)+
  ggtitle("Pclasswise titlewise survival")+
  xlab("title")+
  ylab("No. of people survived/dead")+
  labs(fill="Survived")
```

**We concluded that Master titles in both Pclass1 and Pclass2 have no deaths i.e., survived .**
**so we can conclude that every master. in  Pclass1 and Pclass2 survived in Test dataset observations.**


**Getting a summary of Age Variable **
```{r}
summary(titanic$Age)
```
**Factoring Age Variable to use in plotting**

```{r}
titanic$Age<-as.factor(titanic$Age)
```

**Plotting graph with Age,title,Pclass**

```{r,echo=FALSE}
ggplot(titanic[1:nrow(train),],aes(x=Age,fill=Survived))+
  geom_bar(position="dodge")+
  facet_wrap(~title+Pclass)+
  ggtitle("Pclasswise titlewise survival")+
  xlab("Age")+
  ylab("No. of people survived/dead")+
  labs(fill="Survived")
```

**we concluded that almost every Miss. and Mrs. in Pclass1 and Pclass2 survived**

# Examining the survival rate by sibling/spouse 

**Factoring SibSp variable**

```{r}
titanic$SibSp<-as.factor(titanic$SibSp)
```
**Plotting graph with SibSp,title,Pclass**

```{r,echo=FALSE}
ggplot(titanic[1:nrow(train),],aes(x=SibSp,fill=Survived))+
  geom_bar(position="dodge")+
  facet_wrap(~title+Pclass)+
  ggtitle("Pclasswise titlewise survival")+
  xlab("Siblings/Spouse")+
  ylab("No. of people survived/dead")+
  labs(fill="Survived")
```

**We concluded that every master in Pclass3 with 0,1 siblings/spouse survived**

# Examining  the survival rate by parents/child

**Factoring Parch variable**

```{r}
titanic$Parch<-as.factor(titanic$Parch)
```
**Plotting graph with Parch,title,Pclass**

```{r,echo=FALSE}
ggplot(titanic[1:nrow(train),],aes(x=Parch,fill=Survived))+
  geom_bar(position="dodge")+
  facet_wrap(~title+Pclass)+
  ggtitle("Pclasswise titlewise survival")+
  xlab("Parent/Children")+
  ylab("No. of people survived/dead")+
  labs(fill="Survived")
```


# Examining  the survival rate by Fare

**Factoring Fare variable**

```{r}
titanic$Fare<-as.factor(titanic$Fare)
```

**Plotting graph with Fare,title,Pclass**

```{r,echo=FALSE}
ggplot(titanic[1:nrow(train),],aes(x=Fare,fill=Survived))+
  geom_bar(position="dodge")+
  facet_wrap(~title+Pclass)+
  ggtitle("Pclasswise titlewise survival")+
  xlab("Fare")+
  ylab("No. of people survived/dead")+
  labs(fill="Survived")
```

**We concluded that almost every Mr. in Pclass 2 did not survive**


# Examining the survival rate by Embarked

**Factoring Embarked Variable**

```{r}
titanic$Embarked<-as.factor(titanic$Embarked)
```

**Plotting graph with Embarked,title,Pclass**

```{r,echo=FALSE}
ggplot(titanic[1:nrow(train),],aes(x=Embarked,fill=Survived))+
  geom_bar(position="dodge")+
  facet_wrap(~title+Pclass)+
  ggtitle("Pclasswise titlewise survival")+
  xlab("Embarked")+
  ylab("No. of people survived/dead")+
  labs(fill="Survived")
```

**We concluded that no Mr. from Pclass1 and Pclass2 with Embarked "Q" survived**
**also master from Pclass3 did not survive with Embarked "Q"**







